---
title: "Overcoming Hurdles to Green Living: Insights for Community Climate Action"
subtitle: "Understanding What Stops Sustainable Behaviors and How to Tackle It"
author: 
  - Lexi Knight
thanks: "Code and data are available at: [https://github.com/LexiKnight/toronto_climate/tree/main](https://github.com/LexiKnight/toronto_climate/tree/main)."
date: today
date-format: long
abstract: "This paper explores the factors limiting individuals from adopting environmentally friendly behaviors, drawing on survey data to understand motivations and barriers. Findings reveal that while many people are aware of climate change, obstacles like cost, convenience, and perceived effectiveness hinder sustainable actions. Additionally, the research highlights which methods of communication are most effective across different age groups. By addressing these barriers and tailoring strategies for engagement, cities can better empower residents to contribute to meaningful climate action, fostering a more sustainable urban future."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---


# Introduction

Climate change remains one of the most pressing global challenges, with rising temperatures, extreme weather events, and environmental degradation threatening ecosystems and human well-being. Despite the urgent need for individual and collective action, many people continue to struggle with adopting sustainable behaviors in daily life. Governments and communities are increasingly interested in understanding how to motivate environmentally friendly practices, but significant gaps remain in identifying and addressing the obstacles that prevent action. This paper seeks to bridge that gap by examining not only what prevents individuals from engaging in sustainable behaviors but also what communication strategies can effectively inspire change.

This study focuses on a survey conducted in 2018, which gathered insights into people’s awareness of climate change, their likelihood of taking specific climate-friendly actions, the reasons behind inaction, preferred ways of receiving climate-related information, and demographic characteristics such as age and education. We investigate the extent to which people feel informed about climate change, their readiness to adopt energy-efficient practices, reduce waste, and embrace greener transportation options, among other behaviors. Our analysis also sheds light on the barriers individuals face, ranging from financial and logistical concerns to doubts about the impact of their actions, and explores how demographic factors influence communication preferences.

Addressing the gap between climate awareness and action is critical for effective community climate initiatives. Studies show that while awareness of environmental issues is widespread, behavioral change remains limited (Gifford, 2011). Additionally, research indicates that communication strategies tailored to specific demographics can enhance engagement and effectiveness (Moser & Dilling, 2007). Yet, there is little empirical evidence that explores how personalized communication and targeted interventions can overcome barriers to sustainable behaviors. Our study aims to fill this gap by estimating the causal impact of various factors on individuals’ willingness and ability to engage in environmentally friendly practices and identifying the best strategies for promoting sustainable change.

We found that practical obstacles, such as cost and convenience, alongside perceptions of ineffectiveness, are the primary reasons for inaction. Moreover, age and education level significantly shape how people prefer to receive information about climate action, suggesting that a one-size-fits-all communication strategy may be insufficient. These findings have substantial implications for policymakers and community leaders seeking to design more impactful climate initiatives. By addressing key barriers and tailoring engagement strategies, cities can empower individuals to adopt sustainable behaviors, making meaningful progress toward climate resilience and environmental sustainability.

The remainder of this paper is structured as follows: @sec-data presents the methods used to analyze these demographic differences, followed by @sec-results, which details the results of the comparison between Rosedale-Moore Park and Downtown Yonge East. @sec-discussion offers a discussion of the findings, highlighting the implications for understanding residential segregation in Toronto. Finally, @sec-conclusion concludes with a reflection on the limitations of the study and suggestions for future research directions.

 

# Data {#sec-data}

## Software and R-packages
## Methodology 
### Data Collection
### Data Cleaning
### Data Analysis
## Features
### Age
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-one
#| fig-cap: The chart illustrates the distribution of age for the 2018 survey in the City of Toronto, highlighting the major age categories.

# install libraries 
# install.packages(c("tidyverse", "dplyr", "ggplot2", "here", "stringr", "knitr"))

# Load necessary libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)
library(stringr)
library(knitr)

# Read in the data from your CSV file
analysis_data <- read_csv(here("data/02-analysis_data/analysis_data.csv"))

# Create a histogram to show the distribution of age
age_plot <- ggplot(data, aes(x = age_18)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(
    title = "Age Distribution of Survey Respondents (2018)",
    x = "Age",
    y = "Frequency"
  ) +
  theme_minimal()

# Display the plot
print(age_plot)

# calculate summary statistics 
mean_age <- mean(data$age_18, na.rm = TRUE)
median_age <- median(data$age_18, na.rm = TRUE)
min_age <- min(data$age_18, na.rm = TRUE)
max_age <- max(data$age_18, na.rm = TRUE)
# R does not have built in function thus needs to be defined
sd_age <- sd(data$age_18, na.rm = TRUE)
mode_age <- function(x) {
  unique_x <- unique(x)
  unique_x[which.max(tabulate(match(x, unique_x)))]
}
mode_value <- mode_age(data$age_18)

# Create a data frame with the summary statistics
summary_stats_age_18 <- data.frame(
  Statistic = c("Mean Age", "Median Age", "Mode Age", "Standard Deviation", "Minimum Age", "Maximum Age"),
  Value = c(mean_age, median_age, mode_value, sd_age, min_age, max_age)
)

# Use kable to generate a nice table
kable(summary_stats_age_18, 
      caption = "Summary Statistics for Age of Survey Respondents", 
      col.names = c("Statistic", "Value"),
      align = c("l", "c"))


```

### Highest Level of Education Completed
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-one
#| fig-cap: The chart illustrates the distribution of highest level of education acheived for the 2018 surveyin the city of Toronto, highlighting academic achievement.

# install libraries 
# install.packages(c("tidyverse", "dplyr", "ggplot2", "here", "stringr", "knitr"))

# Load necessary libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)
library(stringr)
library(knitr)

# Read in the data from your CSV file
analysis_data <- read_csv(here("data/02-analysis_data/analysis_data.csv"))

# Create a bar chart for the highest level of education
ggplot(data = analysis_data, aes(x = factor(highest_level_educ_18, 
       levels = c("High school or less",
                  "Some community college, vocational, trade school",
                  "Completed community college, vocational, trade school",
                  "Some university",
                  "Completed undergraduate degree",
                  "Post graduate/professional school", 
                  "Prefer not to answer")))) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(
    title = "Distribution of Highest Level of Education Achievement",
    x = "Level of Education",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.margin = margin(10, 10, 10, 10), # Adds margin around the plot
    plot.title = element_text(hjust = 0.5) # Centers the title
  ) + 
  # Wrap x-axis labels
  scale_x_discrete(labels = function(x) str_wrap(x, width = 20)) + 
  # Y-axis ticks every 25
  scale_y_continuous(breaks = seq(0, max(table(analysis_data$highest_level_educ_18)), by = 25)) 

# Adjust the size of the figure (increase height)
  ggsave("education_chart.png", width = 10, height = 6)
```

### Extent Feel Informed
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-one
#| fig-cap: The bar chart illustrates the distribution of the extent individuals feel informed about what causes climate change in the City of Toronto, based on the 2018 survey.

# install libraries 
# install.packages(c("tidyverse", "dplyr", "ggplot2", "here", "stringr", "knitr"))

# Load necessary libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)
library(stringr)
library(knitr)

# Read in the data from your CSV file
analysis_data <- read_csv(here("data/02-analysis_data/analysis_data.csv"))

# Create a bar chart for the extent people feel informed about climate change
ggplot(data = analysis_data, aes(x = factor(extent_consider_informed_18, 
       levels = c("Extremely informed", 
                  "Very informed", 
                  "Not very informed", 
                  "Not at all informed")))) +
  geom_bar(fill = "lightgreen", color = "black") +
  labs(
    title = "Extent People Feel Informed About Climate Change",
    x = "Level of Informedness",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5) # Center the title
  ) 


```

### Likelihood Taking Action
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-one
#| fig-cap: The bar chart illustrates the distribution of the likelihood of taking certain actions to address climate change in the City of Toronto based on the 2018 survey. 

# install libraries 
# install.packages(c("tidyverse", "dplyr", "ggplot2", "here", "stringr", "knitr"))

# Load necessary libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)
library(stringr)
library(knitr)

# Read in the data from your CSV file
analysis_data <- read_csv(here("data/02-analysis_data/analysis_data.csv"))

# Create reshaped data for plotting
action_data_long <- analysis_data %>%
  pivot_longer(cols = starts_with("likelihood_action"),
               names_to = "Action",
               values_to = "Likelihood") %>%
  mutate(Action = recode(Action,
                         likelihood_action_home_improvement_18 = "Home Improvement",
                         likelihood_action_reduce_hydro_18 = "Reduce Hydro Use",
                         likelihood_action_minimize_car_18 = "Minimize Car Use",
                         likelihood_action_vehicle_electric_18 = "Use Electric Vehicle",
                         likelihood_action_protein_alternative_18 = "Alternative Proteins",
                         likelihood_action_reduce_waste_18 = "Reduce Waste",
                         likelihood_action_green_product_18 = "Green Products",
                         likelihood_action_short_distance_18 = "Short Distance Travel",
                         likelihood_action_sort_waste_18 = "Sort Waste")) %>%
  filter(!is.na(Likelihood)) %>% # Remove rows where Likelihood is NA
  mutate(Likelihood = factor(Likelihood)) # Convert Likelihood to a factor

# Plot the distribution of likelihood of actions to address climate change
ggplot(action_data_long, aes(x = Action, fill = Likelihood)) +
  geom_bar(stat = "count", position = "dodge") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), # Rotate x-axis labels for readability
        axis.title.y = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  labs(
    title = "Likelihood of Taking Action to Address Climate Change in Toronto (2018)",
    x = "Actions to Address Climate Change",
    y = "Count of Responses",
    fill = "Likelihood"
  ) +
  scale_fill_manual(values = c("Already doing this or have done this" = "green",
                               "Very likely" = "blue",
                               "Somewhat likely" = "yellow",
                               "Somewhat unlikely" = "orange",
                               "Very unlikely" = "red"))
```

### Barriers Taking Action
### Best Mode of Delivery

# Model {#sec-model}
The goal of our modelling strategy is twofold. Firstly,...
Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-appendix].
## Model set-up
Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model Specifications ###
### Model Justification 
We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.
#### Response Variable ####
#### Input Variabless ####
#### Model Structure ####
#### Parameter Estimation ####

# Results


# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Limitations

the following categories were not options under reasons unlikely to perform action   "Use public transit more",
  "Cycle more",
  "Walk more",
  "Purchase electric/hybrid vehicle in next 1-3 years",
  but they were categories for which actions are you most likely to take part in to address climate change.
  
no indiivdual data for 2021, used summary statistics instead

## Future Directions

# Conclusion {#sec-conclusion}

\newpage

\appendix

# Appendix {#sec-appendix}


\newpage


# References


